OPTIMAL PRICING AND INFORMAL SHARING: \\ EVIDENCE FROM PIPED WATER IN MANILA

### 2018-12-09 phil_paper/jmp_wjv.tex
matlab est baby steps

%How should public utilities set fixed fees and marginal prices to provide access while covering costs?
%Why are there fixed vs. marginal?
%Public utilities recover fixed and marginal costs through a combination of fixed and marginal %prices.
%WHY CARE?! THE POOR!

	In developing cities, high costs of connecting to piped water lead many households to use low-quality, non-piped sources.  In response, cities often set low fixed prices so that more households connect and high marginal prices to cover production costs.  This paper identifies a common alternative source --- sharing a neighbor's connection --- that has large implications for these pricing strategies.  With microdata from 1.5 million water connections in Manila, I structurally estimate water demand in the presence of water sharing and solve the city's pricing problem.  I find that high marginal prices weaken incentives to share water between neighbors.  In contrast, the optimal pricing policy features high fixed and low marginal prices, increases shared connections, and ensures nearly universal access to piped water compared to current prices in Manila.

% 	However, I find that these prices

% I find that

% 	In particular, I find that high marginal prices weaken incentives to share connections,

% 	access and welfare by weakening incentives to share water.

% 	or share with their neighbors' connections.

% 	.  I then analyze the regulator's pricing problem: choosing fixed and marginal prices to maximize welfare while covering costs.

% 	With these estimates, I find that the presence of sharing reverses

% 	At the same time, cities must choose relative fixed and marginal prices cover production costs by ,

% 	which have large implications for piped water access and usage.  Cities often subsidize

% 	 low fixed and high marginal prices to offset other connection costs and provide access especially for the poor; however, I find that

% 	often leading them to use from low-quality vendors or share with their neighbors' connections.

% 	In developing cities, households face many costs of connecting to piped water, often leading them to use from low-quality vendors or share with their neighbors' connections.  At the same time, many cities must set prices to cover production costs.

% 	Many cities respond by adjusting fixed and marginal prices to ensure access while covering costs.

% 	To ensure piped water access, cities are often constrained to adjusting fixed and marginal prices while covering production costs.

% 	The primary way cities can respond is by  setting fixed and marginal prices still

% 	respond by adjusting the fixed and marginal prices

% 	Since water utilities must often cover their costs, cities are constrained to

% 	With data from 1.5 million water connections in Manila, I structurally estimate household water demand to analyze optimal fixed and marginal prices for piped water in the presence of water sharing.  Cities often set low fixed and high marginal prices to offset other connection costs and provide access especially for the poor; however, I find that

% 	To offset connection costs, cities often subsidize the fixed price with high marginal prices

% 	 to provide access especially for the poor; however,

% 	Using data on 1.5 million water connections in Manila, I structurally estimate household water demand across three sources: purchasing directly from the provider, sharing with a neighbor's tap, or buying from a small-scale vendor.

% to maximize welfare while covering their costs

% 	This paper analyzes how

% 	To offset these connection costs, cities often subsidize their fixed connection fees with high marginal prices to provide access especially for the poor; however,

% 	Public utilities choose fixed and marginal prices to recover costs.

% 	This paper analyzes the welfare effects of pricing policies for piped water in the presence of water sharing.

% 	Public utilities subsidize fixed connection fees with high marginal prices to provide access especially for the poor.

% 	 but this policy can have the opposite effect when many households share connections.

% 	The model predicts that low fixed fees and high marginal prices decrease access and welfare by weakening incentives to share water.

% 	In contrast, the optimal pricing policy features a high fixed fee and low marginal price, increases shared connections, and ensures nearly universal access to piped water compared to current prices in Manila.

% 	I establish that informal sharing networks provide \input{tables/shr_use_from_neighbor}\unskip\% 
% 	of total piped water access.

	%This policy improves welfare by up to \input{tables/tpt_improvement_percent_sharing_group}\unskip\% of consumer surplus.

	% or \input{tables/tpt_improvement_income_sharing_group}\unskip\% of household income

%	When households frequently share connections, this strategy may reduce access by weakening incentives to share.
%weaken incentives for households to extend water access to their neighbors through sharing 
%this strategy may have the opposite effect when connections are frequently shared between households.  
%	I analyze the welfare impacts of pricing for piped water in a context where households may choose to share their water connections with their neighbors.  
%	based on first-world experience

	%This policy improves welfare by up to \input{tables/tpt_improvement_percent_sharing_group}\unskip\% of consumer surplus or \input{tables/tpt_improvement_income_sharing_group}\unskip\% of household income.

@@

These estimates allow me to solve a standard regulator pricing problem: setting prices to cover fixed and marginal costs while maximizing consumer surplus.  I also let the regulator place extra weight on surplus for smaller, poorer users.  In the presence of sharing, the standard approach --- low fixed fees and high marginal prices --- excludes many small-users from piped water by reducing incentives to share connections.  Instead, these households substitute toward non-piped water sources, which have lower marginal prices, but varying quality.\footnote{Common non-piped water sources in Manila include deepwells, tanker trucks, and water refilling stations.}  The opposite approach --- high fixed fees and low marginal prices --- maximizes both access and welfare even for smaller users.  High fixed fees are paid in full by households who connect individually and are split between sharing households.  Informal sharing arrangements enable small-users to access piped water from their neighbors for lower marginal prices than those offered by non-piped sources.  Shared connections also let the city save on installation and maintenance costs, providing an efficient means for extending piped water to low-demand households.\footnote{Cost savings are passed on to consumers through lower prices.}

The welfare impacts of these pricing policies depend not only on household-level demand, but also on the relative costs of sharing water; these may include both marginal costs (ie. time spent fetching from neighbors or reduced pressure from re-piping water) as well as savings from splitting the fixed costs of a connection.  Measuring these features presents two central challenges: how can we recover household-level demand when we only observe water consumption for each connection, and how can we separately identify the costs of sharing and preferences for water?  The empirical approach addresses these challenges in three steps.

The first step builds a tractable model of water sharing to estimate household-level demand from monthly usage data for each connection.  Informed by new survey evidence, this model embeds how (1) households face challenges in observing each other's monthly demand and (2) households often split the monthly water bill evenly.  These assumptions on sharing behavior ensure that connection-level demand can be expressed as a sum of underlying household demands even in the presence of non-linear prices.  I use variation in the non-linear price schedule (both across and within connections) to identify water demand for each connection before disaggregating preferences to the household-level.  This approach extends previous models of optimization over non-linear prices by including sharing and allowing households to anticipate unobserved shocks to their demand (\cite{moffitt1986econometrics,burtless1978effect}).

To estimate the marginal costs of sharing water, the second step analyzes a quasi-experiment where breaks in service pipes force some households to switch from their own water connections to their neighbors' connections.  I then measure how much less water households use from their neighbors' connections and attribute this gap to the marginal costs of sharing.  This exercise recovers sizable marginal costs on the order of \input{tables/hassle_percentage}\unskip\% of the average marginal water price.\footnote{This revealed-preference approach builds on a descriptive literature that identifies potential sources of water sharing frictions (\cite{whittington1992possible,nauges2006water}).}

The third step maps the share of households using from each source --- connecting alone, sharing with a neighbor, or using from a non-piped vendor --- into an estimate of each source's relative fixed cost to consumers.  The empirical approach leverages the assumption that given water preferences (from step one) and marginal sharing costs (from step two), any remaining variation in water source choices must be due to differences in their fixed costs.  Since sharing decisions are strategic in nature, I specify a simple game where households sequentially choose their sources and agree to split any surplus from sharing.

@@

With these estimates, I simulate a series of counterfactual policies.  First, I consider a connection fee subsidy policy consistent with recommendations from the World Bank, NGOs, and other researchers to improve water access for the poor (\cite{komives2005water,jimenez2014factors,mcintosh2003asian}).  I discount the fixed fee by 12\% while increasing the tariff schedule proportionally to ensuring that the water provider exactly covers costs.  This policy actually reduces access by incentivizing households to substitute toward non-piped vendors.  Next, I compute an optimal tariff schedule to maximize consumer surplus while exactly covering production costs.  Compared to the current tariff, I find that a simple, two-part tariff characterized by a fixed fee and low, flat marginal price can produce gains on the order of \input{tables/tpt_improvement_percent_sharing_group}\unskip\% of consumer surplus.\footnote{With quasi-linear utility, consumer surplus for a given tariff is calculated as the difference between household utility under this tariff and household utility from using the non-piped, vendor option.}  These gains are driven both by pricing closer to marginal cost as well as by increasing sharing, which reduces the number of connections installed and maintained by the utility.

@@

%Yet, since sharing involves costly redistribution of water, the optimal tariff subsidizes individual connections on the margin by pricing above marginal cost and setting the fixed fee below the cost of installing and maintaining each connection.

% analyzes a quasi-experiment where households face leaks in their service pipes, forcing them to substitute to a neighbor's connection.

% The model maps the extent to which these households are able to offset their consumption to a neighbor's connection into an estimate of the marginal costs of fetching water.  This exercise recovers sizable sharing costs on the order of \input{tables/hassle_percentage}\unskip\% of the average marginal water price.  This quasi-experimental approach builds on a descriptive literature pointing to potential costs of sharing water (\cite{whittington1992possible,nauges2006water}).

% In Manila, I establish that 23\% of households access water through a neighbor's connection by connecting plumbing directly to the meter (\input{tables/access_pipes}\unskip\%), extending plastic pipes (\input{tables/access_single_tap}\unskip\%), or fetching water with containers (\input{tables/access_fetch}\unskip\%).  Time spent fetching water or reduced pressure-levels from informally re-piping both increase the marginal cost of sharing.  
%--- what is the relative fixed cost of

%combines monthly usage to recover household-level demand.
% recover household-level from connection usage,

% I develop a simple model of the water sharing contract that captures two key features of these relationships informed by survey data:

% The model characterizes monthly consumption choices when households face non-linear budget constraints and imperfect control over their exact consumption levels.

% Greater demand heterogeneity increases the scope for usage-based price discrimination to both redistribute surplus and efficiently raise revenue.

% Sharing costs determine how households can respond by changing their water source alongside their consumption patterns.

%  while low sharing costs allow households

% To evaluate the impacts of different pricing policies, I build a model of household water demand that includes both initial source choices as well as monthly consumption choices.

%For sharing households, I develop a simple model of the water sharing contract that captures two key features of these relationships: (1) households face challenges in observing each other's demand and (2) households often split the monthly water bill evenly.  The model characterizes monthly consumption choices when households face non-linear budget constraints and imperfect control over their exact consumption levels.  This approach extends previous models of optimization over non-linear budget constraints by allowing households to anticipate unobserved shocks to their demand (\cite{moffitt1986econometrics,burtless1978effect}).

% The estimation combines water source and consumption decisions to structurally recover preferences and unobserved costs of water sharing.  To include reduced-form approaches to identification within the context of a structural model, I divide the estimation into three steps.  The first step uses variation in the non-linear price schedule (both across and within households) to identify water demand for each household using monthly consumption data.

%substitute away from sharing arrangements toward both individual connections and
%This policy reduces access by incentivizing households to substitute away from sharing arrangements toward both individual connections and water vendors (which mainly include deepwells and water refilling stations).

%features a much lower fixed fee and higher marginal price and redistributes substantial surplus to low users as compared to the optimal tariff.

%features a slightly lower fixed fee and higher marginal price, produces universal access to piped water and reduces sharing, but reallocates little consumer surplus to low-demand households.

%and increasing marginal prices; it also increases access and efficiently redistributes consumer surplus to low users as compared the optimal tariff.  These results provide evidence that the presence of informal sharing networks constrains pricing policy as an effective tool for redistribution.

%The social tariff features a much higher fixed fee and lower marginal price, does not improve access to piped water, and reallocates little consumer surplus to low-demand households at a large cost to average welfare.  By contrast, in a counterfactual world without sharing, the social tariff features a fixed fee subsidy and increasing marginal prices; it also increases access and efficiently redistributes consumer surplus to low users as compared the optimal tariff.  These results provide evidence that the presence of informal sharing networks constrains pricing policy as an effective tool for redistribution.

@@

% Despite the relevance of water sharing arrangements to pricing policy, previous studies have provided either a descriptive analysis of water sharing (\cite{whittington1992possible}) or 
% in fixed fee incentivizes households to substitute away from water sharing arrangements toward individual connections as well as  , it also increases the share of households using from a local 
%among low-demand households by as much as \input{tables/discount_low_welfare_change}\unskip\% of monthly income.  
%%%% WHAT IS ACCESS DUDE!!!!!!!

%By using from their neighbors, smaller-users also benefit from low marginal prices and avoid sourcing from non-piped water.  
% Non-linear pricing helps policymakers achieve both efficiency (by targeting more inelastic households with higher prices) as well as equity (by targeting poorer households with lower prices).
% excluding many smaller-users from piped water.\footnote{In the model, households can substitute to a vendor option that encompasses all non-piped sources of water, which commonly include deepwells, tanker trucks, and water refilling stations.}  
%The model is designed to match several features from new survey evidence on sharing contracts that I fielded as part of this study.
%In the context of Manila, I find that the most disadvantaged households are also most likely to fetch water from their neighbors.
%In Manila, I establish that 23\% of households access water through a neighbor's connection by connecting plumbing directly to the meter (\input{tables/access_pipes}\unskip\%), extending plastic pipes (\input{tables/access_single_tap}\unskip\%), or fetching water with containers (\input{tables/access_fetch}\unskip\%). 
% Manila 
% To measure these features, I merge administrative data covering 1.5 million water connections in Manila from 2010 to 2015 with detailed survey data on the number of households and people using each water connection.  
% With comprehensive geographic coverage of half of Manila and a variety of price variation, these data allow for a rich characterization of water demand.  Additional variation in water source choices identify the relative fixed and marginal costs associated with sharing water with a neighbor.  
% I find that shared connections allow the city to save on installation and maintenance costs, providing an efficient means for extending piped water to low-demand households at low marginal prices.  
%At the same time, informal sharing networks serve as resale markets for water, undermining the extent to which cities can redistribute surplus through increasing marginal prices.
%  Shared connections record high levels of monthly consumption, which can expose them to high prices when marginal prices increase with usage .

%%% OLD

%%  This strategy has counterintuitive effects in a context with widespread sharing by subsidizing costs for high

%%  Served by a large, government-regulated water monopoly, Manila provides an ideal setting since over \input{tables/shr_1or2hh}\unskip\% of households share water connections.  Given many positive health exernalities from access to safe, reliable water (\cite{galiani2005water,gamper2010impact,kremer2011spring}), redesigning pricing regimes to boost access and usage especially among the poor can have significant welfare consequences.

%%  Over 70\% of developing cities have water pricing regimes characterized by high fixed connection fees and marginal prices that increase with monthly usage to cross-subsidize low-income users (\cite{hoque2013state}).  From a theoretical perspective, the optimal pricing strategy to maximize surplus especially for low-income households while covering costs of production depends on the frictions associated with sharing water.  These sharing frictions may include time costs fetching water from a neighbor as well as difficulties dividing the monthly water bill among sharing households.  If sharing frictions are large, policymakers may want to disincentivize sharing by implementing a low fixed fee and increasing the marginal price schedule.  On the other hand if sharing frictions are small, policymakers may instead use a high fixed fee and alongside a decreasing marginal price schedule in order to encourage sharing networks, set marginal prices closer to marginal costs, and minimize the number of connections that would need to be installed.  This paper provides empirical estimates of these sharing frictions in an effort to guide these pricing decisions.

%%  From a theoretical perspective, the impacts of this pricing regime depend on the frictions associated both with sharing water as well as connecting to the network.  These sharing frictions may include time costs fetching water from a neighbor as well as difficulties dividing the monthly water bill among sharing households.  If sharing frictions are high, policymakers may want to disincentivize sharing by implementing a low fixed fee and increasing marginal price schedule.  COME BACK TO THIS ISSUE: DO I WANT TO GET INTO OPTIMAL SO SOON!? NOT SURE....

%%  While low-cost sharing networks constrain the set of efficient pricing strategies available to the regulator, these networks ultimately improve welfare for low-income households by providing an additional mechanism for these households to access low-cost piped water.

%%  This paper contributes most directly to an expanding literature on estimating demand for public utilities in developing countries, including water in South Africa (\cite{szabo2015value}) and C{\^o}te d'Ivoire (\cite{diakite2009proposal}) and electricity in Colombia (\cite{mcrae2014infrastructure}).  Previous work has focused on how pricing affects monthly consumption decisions.  This paper extends these models by allowing households to also choose their initial water source.  By structurally modeling source choice, this paper also connects to a small reduced-form literature that emphasizes the importance of fixed costs associated with connecting to a piped water network (\cite{devoto2012happiness,jimenez2014factors,kayaga2007costs}).

@@

Observing that ``connection fees are too high, and lump sum upfront payments exclude the poor,'' the Asian Development Bank has argued for steeply increasing water prices and has emphasized that ``the irony of the situation is that the main way to help the poor is to substantially raise tariffs.''\footnote{This quote is from \cite{mcintosh2003asian} pages 40 \& 45.  The authors advocate for a ``three-block system with the lifeline consumption rate at 0-6 m3 (tariff limited to 5\% of household income for the lowest 10th percentile), consumption of 6-20 m3 charged at a rate to recover all financial costs, and consumption of more than 20 m3 charged at a penalty rate equal to about \$1/m3 (to conserve water)'' (\cite{mcintosh2003asian} p.76).}  In this paper, I show that the presence of informal sharing reverses this common intuition among policymakers: high-fixed fees no longer act as a barrier for low-income households when they can split these fees by sharing with their neighbors.  Moreover, low marginal prices help to prevent low-income households from substituting toward non-piped, vendor sources.

@@

Taken together, this paper suggests that households may act as efficient subcontractors in redistributing water to their neighbors.  Instead of sending trained engineers to install and maintain connections in dense neighborhoods, providers can instead rely on informal networks to extend water distribution.  From this perspective, cities may benefit from considering how prices not only might encourage access and redistribute consumer surplus, but also may affect the endogenous formation of water sharing networks.  For example, price discriminating according to neighborhood income instead of by monthly usage levels may achieve similar goals of redistribution without distorting sharing networks.  While this policy may not be appropriate in Manila since neighborhoods are often very heterogeneous in income, this approach may work well in other contexts where neighborhoods are more segregated, such as in Colombia where electricity rates vary according to the social category of each neighborhood (\cite{mcrae2014infrastructure}).

Since sharing networks constrain the menu of pricing policies available to policymakers, cities may benefit from considering a broader range of policies to improve welfare from water use.  For example, cities may be able to lower the frictions associated with these sharing networks by providing technical resources for managing plumbing networks or distributing extra water meters to assist households in accurately splitting the bill.  The empirical results also confirm previous findings that substantial barriers exist for households to connect to the network (\cite{devoto2012happiness}).  Cities may be able to similarly achieve substantial welfare gains by streamlining administrative requirements for new connections and by allowing households to pay connection fees in smaller increments.

>>>>

	In developing cities, high costs of connecting to piped water lead many households to use low-quality, non-piped sources.  In response, cities often set low fixed prices so that more households connect and high marginal prices to cover production costs.  This paper identifies a common alternative source --- sharing a neighbor's connection --- that has large implications for these pricing strategies.  With microdata from 1.5 million water connections in Manila, I structurally estimate water demand in the presence of water sharing and solve the regulator's problem: setting prices to maximize consumer surplus while covering production costs.  I find that low fixed and high marginal prices weaken incentives to share water between neighbors.  In contrast, the optimal pricing policy features high fixed and low marginal prices, increases shared connections, and ensures nearly universal access to piped water compared to current prices in Manila.

@@

These estimates allow me to solve a standard constrained regulator pricing problem: setting prices to cover fixed and marginal costs while maximizing consumer surplus.  I also let the regulator place extra weight on surplus for smaller, poorer users.  In the presence of sharing, the standard approach --- low fixed fees and high marginal prices --- excludes many small-users from piped water by reducing incentives to share connections.  Instead, these households substitute toward non-piped water sources, which are easy to access, have relatively lower marginal prices, and provide varying water quality.\footnote{Common non-piped water sources in Manila include deepwells, tanker trucks, and water refilling stations.  Without fixed prices, land-tenure requirements, or administrative expenses, these sources often have low costs of access.}  The opposite approach --- high fixed fees and low marginal prices --- maximizes both access to piped water and welfare even for smaller users.  High fixed fees are paid in full by households who connect individually and are split between sharing households.  Informal sharing arrangements enable small-users to access piped water from their neighbors for relatively lower marginal prices than those offered by non-piped sources.  By reducing the total number of connections, sharing arrangements also let the city save on installation and maintenance costs, providing an efficient means for extending piped water to low-demand households.\footnote{In this regulatory arrangement, cost savings are passed on to consumers through lower prices.}

The welfare impacts of these pricing policies depend not only on household-level demand, but also on the relative costs of sharing water; these may include both marginal hassle costs (ie. time spent fetching from neighbors or reduced pressure from re-piping water) as well as savings from splitting the fixed costs of a connection.  Measuring these features presents two central challenges: how can we recover household-level demand when we observe water consumption for connections serving multiple households, and how can we separately identify the costs of sharing and preferences for water?  The empirical approach addresses these challenges in three steps.

The first step builds a tractable model of water sharing to estimate household-level demand from monthly usage data for each connection.  Informed by new survey evidence, this model embeds how (1) households face challenges in observing each other's monthly demand and (2) households often split the monthly water bill evenly.  These assumptions on sharing behavior ensure that connection-level demand can be expressed as a simple function of underlying household demands even in the presence of non-linear prices.  I use variation in the non-linear price schedule (both across and within connections) to identify water demand for each connection before disaggregating preferences to the household-level.  This approach extends previous models of optimization over non-linear prices by including sharing and allowing households to anticipate unobserved shocks to their demand (\cite{moffitt1986econometrics,burtless1978effect}).

To estimate the marginal hassle costs of sharing water, the second step analyzes a quasi-experiment where breaks in service pipes force some households to switch from their own water connections to their neighbors' connections.  I then measure how much less water households use from their neighbors' connections and attribute this gap to the marginal hassle costs of sharing.  This exercise recovers sizable marginal hassle costs on the order of \input{tables/hassle_percentage}\unskip\% of the average marginal water price.\footnote{This revealed-preference approach builds on a descriptive literature that identifies potential sources of water sharing frictions (\cite{whittington1992possible,nauges2006water}).}

The third step maps the share of households using from each source --- connecting alone, sharing with a neighbor, or using from a non-piped source --- into an estimate of each source's relative fixed cost to consumers.  The empirical approach leverages the assumption that given water preferences (from step one) and marginal sharing costs (from step two), any remaining variation in water source choices must be due to differences in their fixed costs.  Since sharing decisions are strategic in nature, I specify a simple game where households sequentially choose their sources and agree to split any surplus from sharing.

@@

With these estimates, I simulate a series of counterfactual policies.  First, I consider a connection fee subsidy policy consistent with recommendations from the World Bank, NGOs, and other researchers to improve water access for the poor (\cite{komives2005water,jimenez2014factors,mcintosh2003asian}).  I discount the fixed fee by 12\% while increasing the tariff schedule proportionally to ensuring that the water provider exactly covers costs.  This policy actually reduces access because households stop sharing connections and substitute towards non-piped sources.  Next, I compute an optimal tariff schedule to maximize consumer surplus while exactly covering production costs.  Compared to the current tariff, I find that a simple, two-part tariff characterized by a fixed fee and low, flat marginal price can produce gains on the order of \input{tables/tpt_improvement_percent_sharing_group}\unskip\% of consumer surplus.\footnote{With quasi-linear utility, consumer surplus for a given tariff is calculated as the difference between household utility under this tariff and household utility from using the non-piped, vendor option.}  These gains are driven both by pricing closer to marginal cost as well as by increasing sharing, which reduces the number of connections installed and maintained by the utility.

@@

@@

@@

Observing that ``connection fees are too high, and lump sum upfront payments exclude the poor,'' the Asian Development Bank has argued for cutting upfront payments and steeply increasing water prices, emphasizing that ``the irony of the situation is that the main way to help the poor is to substantially raise tariffs.''\footnote{This quote is from \cite{mcintosh2003asian} pages 40 \& 45.  The authors advocate for a ``three-block system with the lifeline consumption rate at 0-6 m3 (tariff limited to 5\% of household income for the lowest 10th percentile), consumption of 6-20 m3 charged at a rate to recover all financial costs, and consumption of more than 20 m3 charged at a penalty rate equal to about \$1/m3 (to conserve water)'' (\cite{mcintosh2003asian} p.76).}  In this paper, I show that the presence of informal sharing reverses this common intuition among policymakers: high-fixed fees are less of a barrier for low-income households when they can split these fees by sharing with their neighbors.  Moreover, low marginal prices help to prevent low-income households from substituting toward non-piped, vendor sources.

@@

Taken together, this paper suggests that households may act as efficient subcontractors in redistributing water to their neighbors.  Instead of sending trained engineers to install and maintain connections in dense neighborhoods, providers can instead rely on informal networks to extend water distribution.  From this perspective, cities may benefit from considering how prices not only might encourage access and redistribute consumer surplus, but also may affect the endogenous formation of water sharing networks.  For example, price discriminating according to neighborhood income instead of by monthly usage levels may achieve similar goals of redistribution without distorting sharing networks.  While this policy may not be appropriate in Manila since neighborhoods are often very heterogeneous in income, this approach may work well in other contexts where neighborhoods are more segregated.\footnote{For example in Colombia, electricity rates vary according to the social category of each neighborhood (\cite{mcrae2014infrastructure}).}

Since sharing networks make usage-based price discrimination more difficult, cities may benefit from considering a broader range of policies to improve welfare from water use.  For example, cities may be able to lower the frictions associated with these sharing networks by providing technical resources for managing plumbing networks or distributing extra water meters to assist households in accurately splitting the bill.  The empirical results also confirm previous findings that substantial barriers exist for households to connect to the network (\cite{devoto2012happiness}).  Cities may be able to similarly achieve substantial welfare gains by streamlining administrative requirements for new connections and by allowing households to pay connection fees in smaller increments.
